---
title: "EC325 Preliminary Results"
author: "Alex Menr, Tucker Reinhard, and Othar Zaldastani II"
date: "April 24, 2026"
format:
  html:
    pdf-engine: pdflatex
    geometry: "margin=1in"
    df-print: paged
    toc: false
    keep-tex: true
    fig-cap-location: bottom
  
execute:
  echo: false
  warning: false
  cache: true
  freeze: auto

editor:
  mode: source
knitr:
  opts_chunk:
    render: knitr::normal_print
---

```{r}
#| label: setup
#| echo: true
#| results: hide
#| message: false
#| warning: false
#| include: false
# Setup Chunk
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  results = 'hold',
  fig.align = "center",
  attr.source = '.numberLines',
  cache = FALSE
)
library(tidyverse)              # Packages
library(psych)
library(psychTools)
library(dplyr)
library(knitr)
library(kableExtra)
library(epitools)
library(broom)
library(modelsummary)
library(kableExtra)
library(plm)
library(fixest)
set.seed(24)                    # Seed for reproducibility
options(                        # Custom Model Summary
  modelsummary_stars = TRUE,
  modelsummary_output = "markdown",
  modelsummary_shape = model ~ term,
  modelsummary_fmt = 2,
  modelsummary_gof_omit = "")
ms <- function(models, ...) {
  model_name <- deparse(substitute(models))
  g <- broom::glance(models)
  modelsummary(models, stars = FALSE, output = "markdown",
               shape = model ~ term, fmt = 2,
               statistic = c("({std.error})", "p = {p.value}"),
               notes = paste0("Model: ", model_name,
                              " | R² = ", round(g$r.squared, 3),
                              ", Adj. R² = ", round(g$adj.r.squared, 3),
                              ", F = ", round(g$statistic, 2),
                              ", p = ", format.pval(g$p.value, digits = 3),
                              ", N = ", g$nobs), ...)
}
kable.dark <- function(x, ...) {    # Custom kable for debugging in dark mode
  kableExtra::kable_styling(
    knitr::kable(x, ...))}

# Need fig.cap = "Name" for appendix figures
# No fig.cap will make it knit in place
```

```{r load-reshape}
FDI_Data4 <- read.csv("~/Downloads/P_Data_Extract_From_World_Development_Indicators-2/e24a155d-2435-4e2c-954e-222fbe9ccf40_Data.csv")
raw <- FDI_Data4
names(raw) <- gsub("\\.", " ", names(raw)) |> trimws()

long <- raw %>%
  pivot_longer(
    cols      = matches("^X\\d{4}"),
    names_to  = "year_label",
    values_to = "value"
  ) %>%
  mutate(
    year  = as.integer(str_extract(year_label, "\\d{4}")),
    value = suppressWarnings(as.numeric(value)),
    var   = case_when(
      `Series Name` == "Total natural resources rents (% of GDP)"                  ~ "nat_resources",
      `Series Name` == "Control of Corruption: Estimate"                           ~ "corruption",
      `Series Name` == "Government Effectiveness: Estimate"                        ~ "gov_effectiveness",
      `Series Name` == "Political Stability and Absence of Violence/Terrorism: Estimate" ~ "pol_stability",
      `Series Name` == "Gini index"                                                ~ "gini",
      `Series Name` == "Tax revenue (% of GDP)"                                    ~ "tax_revenue",
      `Series Name` == "Expense (% of GDP)"                                        ~ "gov_expense",
      `Series Name` == "Rule of Law: Estimate"                                     ~ "rule_of_law",
      `Series Name` == "Regulatory Quality: Estimate"                              ~ "reg_quality",
      `Series Name` == "Literacy rate, adult total (% of people ages 15 and above)" ~ "literacy",
      grepl("Human capital", `Series Name`)                                        ~ "hci",
      `Series Name` == "Energy use (kg of oil equivalent per capita)"              ~ "energy_use",
      `Series Name` == "Trade (% of GDP)"                                          ~ "trade",
      `Series Name` == "GDP per capita (constant 2015 US$)"                        ~ "gdp_pc",
      `Series Name` == "Foreign direct investment, net inflows (% of GDP)"         ~ "fdi",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(var)) %>%
  select(`Country Name`, `Country Code`, year, var, value)

panel <- long %>%
  pivot_wider(
    id_cols     = c(`Country Name`, `Country Code`, year),
    names_from  = var,
    values_from = value,
    values_fn   = mean
  ) %>%
  rename(country = `Country Name`, iso3 = `Country Code`) %>%
  rename_with(~ gsub("[^a-zA-Z0-9_]", "_", .x)) %>%
  filter(year >= 1996, year <= 2023)
```


```{r}
# Creating Lagged GDP_pc
panel <- panel |> 
  arrange(iso3, year) |> 
  group_by(iso3) |> 
  mutate(gdp3 = dplyr::lead(gdp_pc, 3)) |> 
  ungroup()

# GDP Growth
panel <- panel |> 
  arrange(iso3, year) |> 
  group_by(iso3) |> 
  mutate(gdp_growth = 1-(gdp3/gdp_pc)) |> 
  ungroup()

# FDI Per Capita
panel$fdi.pp = panel$gdp_pc * panel$fdi / 100

# FDI Factor
panel <- panel |> 
  mutate(fdi_pos = as.integer(fdi > 0))

panel.pos <- panel |> filter(fdi > 0)

```

```{r}
# Bivariate Model
bivariate.model = lm(gdp3 ~ fdi, data = panel)
#ms(bivariate.model)

# Bivariate Factor Model
bivariate.model2 = lm(gdp_growth ~ fdi_pos, data = panel)
#ms(bivariate.model2)

# Bivariate Factor Model 2
bivariate.model3 = lm(gdp_growth ~ fdi, data = panel)
#ms(bivariate.model3)
```

```{r}
# Full Model
full <- lm(gdp3 ~ . - gdp_pc - country - year - iso3, data = panel)


# Manual Stepwise Regression
# Removing Variable With Highest p-value
m2 = update(full, . ~ . - gov_effectiveness)

m3 = update(m2, . ~ . - gini)

m4 = update(m3, . ~ . - rule_of_law)

m5 = update(m4, . ~ . - pol_stability)

m6 = update(m5, . ~ . - literacy)

m7 = update(m6, . ~ . - reg_quality)
```
```{r}
#| include: false
df <- panel |> select(-country, -iso3, -year, -gdp_pc, -gini) |> na.omit()

null <- lm(gdp3 ~ 1, data = df)
full <- lm(gdp3 ~ ., data = df)

forward <- step(null, 
                scope = list(lower = null, upper = full),
                direction = "forward")

```


# Bivariate Models
## Primary Bivariate Model 

This model is a bivariate regression of GDP per capita looking 3 years forward and FDI- both in dollar amounts. The coefficient is statistically significant and it displays at all confidence levels though. We find a one dollar increase in FDI correlates with a \$79.33 increase in the GDP per capita ceteris paribus. 
```{r}
ms(bivariate.model)
```


## Other Bivariate Models
Below are some of the other bivariate models we tested but did not find as useful as the primary one.
```{r}
ms(bivariate.model2)
ms(bivariate.model3)
```



# Multivariate Models
## Full Model

```{r}
ms(full)
```

This table displays the full model with each of the variables we intend to control for. A 1 dollar increase in an FDI correlates with a \$62 increase in GDP per capita ceteris paribus. Notably there was a slight decrease in the coefficient for FDI but it is still significantly significant with a p value of .003. We will likely remove some of the government variables (`gov_effectiveness`, `pol_stability`, `corruption`) as they are colinear. Notably this model has an adjusted $R^2$ value of .8 which is far higher than the bivariate adj $R^2$ of .03. 


## Manual Stepwise Model

To create the stepwise model, we start with a full model with every predictor variable. Unsurprisingly, there are a lot of multicollinearity and overplotting. Using a combination of logic and data we choose one variable at a time to remove and then tested the new model without that variable. We continue this process a few more times, resulting in a model with what we believed to be the most important variables. We also use adjusted $R^2$ to compare the models.

```{r}
ms(m7)
```

## Automated Stepwise Model

Alternatively, we also had R  run a forward stepwise model. This method has R start with a null model and continues adding variables one at a time. R tests each model using strictly statistical indicators to evaluate each model. Once the model starts to get worse statistically, R stops adding variables leaving you with the “best” statistical model. This was the resulting model.


```{r}
ms(forward)
```









\pagebreak

# Appendix

```{r}
#| ref.label: !expr knitr::all_labels()
#| echo: true
#| eval: false
```

\vfill

Attributions: Template code comes from `alexpghayes`, modified by Claude and myself.
