---
title: "FDI & GDP Prediction Analysis"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    code_folding: hide
    fig_width: 9
    fig_height: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo      = TRUE,
  warning   = FALSE,
  message   = FALSE,
  fig.align = "center"
)
```

```{r libraries}
library(tidyverse)
library(plm)
library(lmtest)
library(sandwich)
library(xgboost)
library(Metrics)
library(corrplot)
library(scales)
library(zoo)
library(knitr)
library(kableExtra)
```

---

# 1. Load & Reshape

The raw data is in wide format — one column per year. We pivot to long, recode series names, then pivot back to get one row per country-year with each indicator as its own column.

```{r load-reshape}
FDI_Data4 <- read.csv("~/Downloads/P_Data_Extract_From_World_Development_Indicators-2/e24a155d-2435-4e2c-954e-222fbe9ccf40_Data.csv")
raw <- FDI_Data4
names(raw) <- gsub("\\.", " ", names(raw)) |> trimws()

long <- raw %>%
  pivot_longer(
    cols      = matches("^X\\d{4}"),
    names_to  = "year_label",
    values_to = "value"
  ) %>%
  mutate(
    year  = as.integer(str_extract(year_label, "\\d{4}")),
    value = suppressWarnings(as.numeric(value)),
    var   = case_when(
      `Series Name` == "Total natural resources rents (% of GDP)"                 ~ "nat_resources",
      `Series Name` == "Control of Corruption: Estimate"                           ~ "corruption",
      `Series Name` == "Government Effectiveness: Estimate"                        ~ "gov_effectiveness",
      `Series Name` == "Political Stability and Absence of Violence/Terrorism: Estimate" ~ "pol_stability",
      `Series Name` == "Gini index"                                                ~ "gini",
      `Series Name` == "Tax revenue (% of GDP)"                                    ~ "tax_revenue",
      `Series Name` == "Expense (% of GDP)"                                        ~ "gov_expense",
      `Series Name` == "Rule of Law: Estimate"                                     ~ "rule_of_law",
      `Series Name` == "Regulatory Quality: Estimate"                              ~ "reg_quality",
      `Series Name` == "Literacy rate, adult total (% of people ages 15 and above)" ~ "literacy",
      grepl("Human capital", `Series Name`)                                        ~ "hci",
      `Series Name` == "Energy use (kg of oil equivalent per capita)"              ~ "energy_use",
      `Series Name` == "Trade (% of GDP)"                                          ~ "trade",
      `Series Name` == "GDP per capita (constant 2015 US$)"                        ~ "gdp_pc",
      `Series Name` == "Foreign direct investment, net inflows (% of GDP)"         ~ "fdi",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(var)) %>%
  select(`Country Name`, `Country Code`, year, var, value)

panel <- long %>%
  pivot_wider(
    id_cols     = c(`Country Name`, `Country Code`, year),
    names_from  = var,
    values_from = value,
    values_fn   = mean
  ) %>%
  rename(country = `Country Name`, iso3 = `Country Code`) %>%
  rename_with(~ gsub("[^a-zA-Z0-9_]", "_", .x)) %>%
  filter(year >= 1996, year <= 2023)

glimpse(panel)
```

---

# 2. Feature Engineering

```{r features}
panel <- panel %>%
  arrange(iso3, year) %>%
  group_by(iso3) %>%
  mutate(
    # Log GDP — stabilises variance for modelling
    log_gdp_pc    = log(gdp_pc),

    # GDP growth rate
    gdp_growth    = (gdp_pc - lag(gdp_pc)) / lag(gdp_pc) * 100,

    # Lagged GDP — autoregressive signal
    gdp_lag1      = lag(gdp_pc, 1),
    gdp_lag3      = lag(gdp_pc, 3),

    # Lagged FDI — investment effects on GDP are delayed
    fdi_lag1      = lag(fdi, 1),
    fdi_lag3      = lag(fdi, 3),

    # Governance composite — mean of 5 WB indicators
    gov_composite = rowMeans(cbind(corruption, gov_effectiveness,
                                   pol_stability, rule_of_law,
                                   reg_quality), na.rm = TRUE),

    # Rolling 3-year governance — smooths year-to-year noise
    gov_roll3     = rollmeanr(gov_composite, 3, fill = NA),

    # FDI x governance — FDI matters more in high-governance environments
    fdi_x_gov     = fdi_lag1 * gov_composite,

    # Trade
    trade_lag1    = lag(trade, 1),
    trade_growth  = (trade - lag(trade)) / lag(trade) * 100,

    # Resource dependency lag
    nat_res_lag1  = lag(nat_resources, 1),

    # Tax burden change
    tax_change    = tax_revenue - lag(tax_revenue)
  ) %>%
  ungroup()
```

| Feature | Description |
|---|---|
| `log_gdp_pc` | Log GDP per capita — stabilises variance |
| `gdp_lag1/3` | Lagged GDP — autoregressive predictor |
| `fdi_lag1/3` | Lagged FDI — investment effects are delayed |
| `gov_composite` | Mean of 5 World Bank governance indicators |
| `gov_roll3` | 3-year rolling governance mean |
| `fdi_x_gov` | FDI x governance interaction |
| `trade_lag1` | Lagged trade openness |

---

# 3. Exploratory Data Analysis

## 3.1 Data Coverage

Red = sparse, green = good coverage.

```{r coverage-heatmap, fig.height=5}
coverage <- long %>%
  filter(year >= 1960, year <= 2023) %>%
  group_by(var, year) %>%
  summarise(pct_available = mean(!is.na(value)) * 100, .groups = "drop")

ggplot(coverage, aes(x = year, y = var, fill = pct_available)) +
  geom_tile(colour = "white", linewidth = 0.3) +
  scale_fill_gradient2(
    low      = "#d73027", mid = "#fee090", high = "#1a9850",
    midpoint = 50, name = "% countries\nwith data"
  ) +
  labs(title = "Data Coverage by Indicator and Year", x = NULL, y = NULL) +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 3.2 FDI vs Log GDP per Capita

Colour shows governance quality — higher governance countries tend to attract more productive FDI.

```{r fdi-gdp-scatter}
panel %>%
  filter(year == 2015, !is.na(fdi), !is.na(gdp_pc)) %>%
  ggplot(aes(x = fdi, y = log_gdp_pc, colour = gov_composite)) +
  geom_point(alpha = 0.7, size = 2.5) +
  geom_smooth(method = "lm", se = TRUE, colour = "black", linewidth = 0.7) +
  scale_colour_gradient2(
    low      = "#d73027", mid = "#fee090", high = "#1a9850",
    midpoint = 0, name = "Governance\ncomposite"
  ) +
  scale_x_continuous(limits = c(-5, 30)) +
  labs(
    title = "FDI vs Log GDP per Capita (2015)",
    x     = "FDI net inflows (% of GDP)",
    y     = "Log GDP per capita (constant 2015 USD)"
  ) +
  theme_minimal()
```

## 3.3 GDP per Capita Distribution Over Time

```{r gdp-boxplot}
panel %>%
  filter(!is.na(gdp_pc), year %in% seq(1960, 2023, by = 3)) %>%
  mutate(year = factor(year)) %>%
  ggplot(aes(x = year, y = log_gdp_pc)) +
  geom_boxplot(fill = "#2166ac", alpha = 0.6, outlier.size = 0.8) +
  labs(
    title = "Distribution of Log GDP per Capita Over Time",
    x     = NULL,
    y     = "Log GDP per capita (constant 2015 USD)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 3.4 Global FDI Trend

```{r fdi-trend}
panel %>%
  group_by(year) %>%
  summarise(mean_fdi = mean(fdi, na.rm = TRUE),
            sd_fdi   = sd(fdi, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = mean_fdi)) +
  geom_ribbon(aes(ymin = mean_fdi - sd_fdi, ymax = mean_fdi + sd_fdi),
              alpha = 0.2, fill = "#2166ac") +
  geom_line(colour = "#2166ac", linewidth = 1) +
  labs(
    title    = "Global Average FDI Inflows Over Time",
    subtitle = "Shaded band = +/-1 SD across countries",
    x        = NULL,
    y        = "FDI net inflows (% of GDP)"
  ) +
  theme_minimal()
```

## 3.5 Correlation Matrix

```{r corrplot, fig.height=6}
cor_vars <- panel %>%
  select(log_gdp_pc, fdi, fdi_lag1, trade, nat_resources,
         gov_composite, literacy, energy_use, tax_revenue) %>%
  drop_na()

corrplot(
  cor(cor_vars),
  method = "color",
  type   = "upper",
  tl.cex = 0.85,
  col    = colorRampPalette(c("#d73027", "white", "#1a9850"))(200),
  title  = "Correlation Matrix - Key Variables",
  mar    = c(0, 0, 2, 0)
)
```

## 3.6 GDP Trends — Selected Countries

```{r gdp-trends}
showcase <- c("USA", "CHN", "IND", "BRA", "DEU", "NGA", "BGD")

panel %>%
  filter(iso3 %in% showcase, !is.na(gdp_pc)) %>%
  ggplot(aes(x = year, y = gdp_pc, colour = iso3)) +
  geom_line(linewidth = 0.9) +
  scale_y_continuous(labels = comma) +
  labs(
    title  = "GDP per Capita Over Time - Selected Countries",
    x      = NULL,
    y      = "GDP per capita (constant 2015 USD)",
    colour = "Country"
  ) +
  theme_minimal()
```

---

# 4. Panel Regression (Fixed Effects)

Two-way fixed effects control for time-invariant country differences and global year shocks. Standard errors are clustered by country.

```{r panel-regression}
model_data <- panel %>%
  select(country, iso3, year, log_gdp_pc, fdi_lag1, fdi_lag3,
         fdi_x_gov, trade, nat_resources, gov_composite,
         trade_lag1, literacy, energy_use, tax_revenue, gov_expense) %>%
  drop_na() %>%
  filter(is.finite(log_gdp_pc))

pdata <- pdata.frame(model_data, index = c("iso3", "year"))

# Model 1: FDI only
fe1 <- plm(log_gdp_pc ~ fdi_lag1,
           data = pdata, model = "within", effect = "twoways")

# Model 2: FDI + macro controls
fe2 <- plm(log_gdp_pc ~ fdi_lag1 + trade + nat_resources +
             tax_revenue + gov_expense,
           data = pdata, model = "within", effect = "twoways")

# Model 3: Full specification
fe3 <- plm(log_gdp_pc ~ fdi_lag1 + fdi_lag3 + fdi_x_gov +
             gov_composite + trade + trade_lag1 + nat_resources +
             tax_revenue + energy_use + literacy,
           data = pdata, model = "within", effect = "twoways")

robust_se <- function(model) {
  coeftest(model, vcov = vcovHC(model, type = "HC1", cluster = "group"))
}
```

## Model 1 - FDI Only

```{r fe1}
robust_se(fe1)
```

## Model 2 - FDI + Macro Controls

```{r fe2}
robust_se(fe2)
```

## Model 3 - Full Specification

```{r fe3}
robust_se(fe3)
```

## Model Fit Summary

```{r r-squared}
tibble(
  Model       = c("1: FDI only", "2: + Macro", "3: Full"),
  `Within R2` = c(
    round(summary(fe1)$r.squared["rsq"], 4),
    round(summary(fe2)$r.squared["rsq"], 4),
    round(summary(fe3)$r.squared["rsq"], 4)
  )
) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

# 5. Machine Learning - XGBoost

Time-aware split to prevent data leakage: train on 1960-2015, validate on 2016-2019, test on 2020+.

```{r xgboost-setup}
features <- c("fdi_lag1", "fdi_lag3", "fdi_x_gov",
              "trade", "trade_lag1", "nat_resources", "gov_composite",
              "gov_roll3", "literacy", "energy_use",
              "tax_revenue", "gov_expense")

ml_data <- panel %>%
  select(year, log_gdp_pc, all_of(features)) %>%
  drop_na() %>%
  filter(is.finite(log_gdp_pc))

train <- ml_data %>% filter(year <= 2015)
valid <- ml_data %>% filter(year > 2015, year <= 2019)
test  <- ml_data %>% filter(year >= 2020)

make_matrix <- function(df) as.matrix(df %>% select(all_of(features)))

dtrain <- xgb.DMatrix(make_matrix(train), label = train$log_gdp_pc)
dvalid <- xgb.DMatrix(make_matrix(valid), label = valid$log_gdp_pc)
dtest  <- xgb.DMatrix(make_matrix(test),  label = test$log_gdp_pc)
```

```{r xgboost-train}
params <- list(
  objective        = "reg:squarederror",
  eta              = 0.05,
  max_depth        = 4,
  subsample        = 0.8,
  colsample_bytree = 0.8,
  min_child_weight = 5
)

set.seed(42)
xgb_model <- xgb.train(
  params                = params,
  data                  = dtrain,
  nrounds               = 500,
  watchlist             = list(train = dtrain, valid = dvalid),
  early_stopping_rounds = 30,
  print_every_n         = 100,
  verbose               = 1
)
```

## Performance

```{r xgb-performance}
pred_valid <- predict(xgb_model, dvalid)
pred_test  <- predict(xgb_model, dtest)

tibble(
  Split = c("Validation (2016-2019)", "Test (2020+)"),
  RMSE  = c(round(rmse(valid$log_gdp_pc, pred_valid), 4),
             round(rmse(test$log_gdp_pc,  pred_test),  4)),
  MAE   = c(round(mae(valid$log_gdp_pc,  pred_valid), 4),
             round(mae(test$log_gdp_pc,   pred_test),  4))
) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Feature Importance

```{r xgb-importance, fig.height=5}
importance <- xgb.importance(feature_names = features, model = xgb_model)
xgb.plot.importance(importance,
                    main = "XGBoost Feature Importance - GDP Prediction",
                    col  = "#2166ac")
```

## Actual vs Predicted

```{r xgb-scatter}
tibble(
  actual    = valid$log_gdp_pc,
  predicted = pred_valid
) %>%
  ggplot(aes(x = actual, y = predicted)) +
  geom_point(alpha = 0.5, colour = "#2166ac", size = 2) +
  geom_abline(slope = 1, intercept = 0, colour = "red",
              linetype = "dashed", linewidth = 0.8) +
  labs(
    title    = "Actual vs Predicted Log GDP per Capita",
    subtitle = "Validation set (2016-2019) — dashed line = perfect fit",
    x        = "Actual",
    y        = "Predicted"
  ) +
  theme_minimal()
```

---

# 6. Forecast 2024-2030

Each country's predictors held at their 2023 values. Missing values imputed with the global median.

```{r forecast}
latest <- panel %>%
  filter(year == 2023) %>%
  select(iso3, country, all_of(features))

for (v in features) {
  med <- median(latest[[v]], na.rm = TRUE)
  latest[[v]][is.na(latest[[v]])] <- med
}

forecast_rows <- expand_grid(iso3 = latest$iso3, year = 2024:2030) %>%
  left_join(latest, by = "iso3")

dforecast <- xgb.DMatrix(as.matrix(forecast_rows %>% select(all_of(features))))
forecast_rows$pred_log_gdp <- predict(xgb_model, dforecast)
forecast_rows$pred_gdp_pc  <- exp(forecast_rows$pred_log_gdp)
```

```{r forecast-plot}
historical <- panel %>%
  filter(iso3 %in% showcase, !is.na(gdp_pc)) %>%
  select(iso3, year, gdp_pc) %>%
  mutate(type = "Historical")

projected <- forecast_rows %>%
  filter(iso3 %in% showcase) %>%
  select(iso3, year, gdp_pc = pred_gdp_pc) %>%
  mutate(type = "Forecast")

bind_rows(historical, projected) %>%
  ggplot(aes(x = year, y = gdp_pc, colour = iso3,
             linetype = type, group = interaction(iso3, type))) +
  geom_line(linewidth = 0.9) +
  scale_y_continuous(labels = comma) +
  scale_linetype_manual(values = c("Historical" = "solid", "Forecast" = "dashed")) +
  labs(
    title    = "GDP per Capita: Historical + Forecast (2024-2030)",
    subtitle = "XGBoost projection | predictors held at 2023 levels",
    x        = NULL,
    y        = "GDP per capita (constant 2015 USD)",
    colour   = "Country",
    linetype = NULL
  ) +
  theme_minimal()
```

## Top 10 Countries by Projected GDP per Capita in 2030

```{r top10}
forecast_rows %>%
  filter(year == 2030) %>%
  arrange(desc(pred_gdp_pc)) %>%
  select(Country = country, `Projected GDP pc (2030)` = pred_gdp_pc) %>%
  head(10) %>%
  mutate(`Projected GDP pc (2030)` = dollar(round(`Projected GDP pc (2030)`))) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

---

# 7. Export Results

```{r export}
write.csv(panel,         "FDI_GDP_panel_clean.csv",        row.names = FALSE)
write.csv(forecast_rows, "FDI_GDP_forecast_2024_2030.csv", row.names = FALSE)

cat("Saved: FDI_GDP_panel_clean.csv and FDI_GDP_forecast_2024_2030.csv")
```

---

*Analysis produced with R `r getRversion()`. Knit date: `r Sys.Date()`.*
