---
title: "Untitled"
format: html
---

---
title: "SC306 Homework 6"
author: "Othar Zaldastani II"
date: "April 8, 2026"
format:
  pdf:
    df-print: paged
    include-in-header:
      text: |
        \usepackage{xcolor}
        \makeatletter
        \makeatother
        \usepackage{float}
        \usepackage{endnotes}
        \renewcommand{\CommentTok}[1]{\textcolor[rgb]{0.60,0.20,0.20}{#1}}
  docx: default
---

```{r}
#| label: setup
#| echo: true
#| results: hide
#| message: false
#| warning: false
#| include: false

# Setup Chunk
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  attr.source = '.numberLines',
  cache = FALSE)

library(tidyverse)
library(psych)
library(psychTools)
library(dplyr)
library(knitr)
library(kableExtra)
library(epitools)
library(broom)
library(modelsummary)
library(kableExtra)
set.seed(24)

options(
  modelsummary_stars = TRUE,
  modelsummary_output = "markdown",
  modelsummary_shape = model ~ term,
  modelsummary_fmt = 2,
  modelsummary_gof_omit = "")

ms <- function(models, ...) {
  g <- broom::glance(models)
  modelsummary(models, stars = F, output = "markdown",
               shape = model ~ term, fmt = 2,
               statistic = c("({std.error})", "p = {p.value}"),
               notes = paste0("R² = ", round(g$r.squared, 3),
                              ", Adj. R² = ", round(g$adj.r.squared, 3),
                              ", F = ", round(g$statistic, 2),
                              ", p = ", format.pval(g$p.value, digits = 3),
                              ", N = ", g$nobs), ...)}

kable2 <- function(x, ...) {
  kableExtra::kable_styling(
    knitr::kable(x, ...))}

fig_paths  <- character(0)
fig_labels <- character(0)
fig_num    <- 0L

knitr::knit_hooks$set(plot = function(x, options) {
  fig_num    <<- fig_num + 1L
  fig_paths  <<- c(fig_paths, x)
  fig_labels <<- c(fig_labels, options$label)
  paste0("\n\n\\textbf{Figure ", fig_num, "} --- see Figure Appendix.\n\n")})
```


# Question 1
## Part A
```{r Question 1A}
pairs.panels(mtcars, main = "Panels for Q1A")
```

#### Part B
Model 1
```{r Question 1B}
m1 <- lm(mpg ~ hp, data = mtcars)
```

```{r Question 1B Summary, out.width="6in"}
ms(m1)
```

\pagebreak
# Question 2
#### Part A
Plot 2
```{r Question 2A}
plot(mtcars$mpg, mtcars$hp, data = mtcars, main = "Panels for Q2A")
```




\pagebreak
# Figure Appendix

```{r}
#| label: figure-appendix
#| echo: false
#| results: asis

for (i in seq_along(fig_paths)) {
  cat(paste0(
    "```{=latex}\n",
    "\\begin{figure}[H]\n",
    "\\centering\n",
    "\\includegraphics[width=0.8\\linewidth]{", fig_paths[i], "}\n",
    "\\begin{center}\\textbf{Figure ", i, " (", fig_labels[i], ")}\\end{center}\n",
    "\\end{figure}\n",
    "```\n\n"))}
```



\pagebreak
# Code Appendix

```{r}
#| ref.label: !expr knitr::all_labels()
#| echo: true
#| eval: false
```

\vfill

Attributions: I have used some template/styling code from GitHub user alexpghayes.
